Paper Title*

### 2018-12-25 paper.tex
lorawan definition

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands. Focus on this work is one of the most promising LPWAN technologies: Lora. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS). LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. SF assignment of nodes has strong effect on collisions thus network performance. In this work, we implemented a simulation environment to study different SF assignment schemes and we proposed a machine learning technique to optimize SF assignment. Finally, we illustrate simulation results for proposed machine learning SF assignment technique.
\end{abstract}

@@

\par There are several emerging LPWAN technologies. LoRa, SigFox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost. (CITE) Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. (CITE) LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access. (CITE) LoRaWAN networks can be deployed as private network. However, SigFox and NB-IoT supports only operator contracted deployments. (CITE) Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. (CITE) However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. (CITE)
Sigfox maximum data rate is also very low 100 bps, LoRa 50 kbps, NB-IoT 200 kbps.\\
LoRa can change data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between sensitivity and data rate. LoRa has 7 different spreading factor option. Different spreading factor communications are orthogonal to each other.

\par TODO: SF Assignment Issue on LoRa

\par TODO: Proposal for SF Assignment Issue on LoRa

\par TODO: Organization of the paper

\section{LoRa}
\par LoRa is the name of the physical layer Radio/Chipset technology that provides wireless link for low power wide area networks. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). LoRa is registered trademark of Semtech Corporation. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently and incorrectly used for each other.

\subsection{LoRaWAN}
\par LoRaWAN is a media access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to promoting the interoperability and standardization of LoRaWAN. LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides interoperability between different networks. LoRaWAN uses pure ALOHA to access medium. LoRaWAN provides lightweight but powerful standard for IoT applications.

\par TODO: explain end node, gateway, class A

\subsection{LoRa Modulation}
\par TODO

\subsection{Spreading Factor}
\par LoRa CSS modulation can spread the symbols.\\
The spread factor is the ratio between symbol rate and chip rate. The ratio between symbol and chip rate is $2$\textsuperscript{spreadfactor}.\\
Six different spread factors are available (between 7 and 12); increasing the spread factor makes the signal more robust to noise, but decreases the data rate.\\

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora0}
\caption{Different SF ranges.}
\label{fig:problem1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora1}
\caption{Collision between nodes near the GW.}
\label{fig:problem2}
\end{figure}

\section{Other Related Works}
The literature related to the work presented in this paper has started growing recently. LPWAN technologies and especially LoRa/LoRaWAN attract researchers attention lately. We listed some of these works which studying  LoRaWAN and LoRa spreading factor.

@@

\section{Proposed Technique}
\par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF. Figure~\ref{fig:proposal_single_gw}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora2}
\caption{Collision between nodes near the GW can be prevented by using higher SF.}
\label{fig:proposal_single_gw}
\end{figure}

@@

\section{Simulation Environment and Results}
A discrete event simulator is implemented in Python to simulate LoRa network performance. 
https://github.com/tugrulyatagan/simlorafs

@@

\section{Conclusion}
TODO: Summary, future works
\cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}

>>>>

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands. Focus on this work is one of the most promising LPWAN technologies: Lora. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS). LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Spreading factor (SF) assignment of nodes has strong effect on collisions thus network performance. In this work, we implemented a simulation environment to study different SF assignment schemes and we proposed a machine learning technique to optimize SF assignment. Finally, we illustrate simulation results for proposed machine learning SF assignment technique.
\end{abstract}

@@

\par There are several emerging LPWAN technologies. LoRa, SigFox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost. (CITE) Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. (CITE) LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access. (CITE) LoRaWAN networks can be deployed as private network. However, SigFox and NB-IoT supports only operator contracted deployments. (CITE) Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. (CITE) However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. (CITE) Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively. (CITE)

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission air time. (CITE) LoRa supports 6 different spreading factor option. Simultaneous same spreading factor transmissions are prone to collision however different spreading factor transmissions in same channel are orthogonal to each other, thus spreading factor assignment is crucial for overall network performance. (CITE) In this work, we introduce a custom made LoRa discrete event simulator to study LoRa spreading factor assignment issue. We also proposed an spreading factor assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.

\par The following section gives an overview about LoRa and LoRaWAN. Section \ref{Other Related Works} describes other related works. Section \ref{Proposed Technique} describes proposed technique for spreading factor assignment. Simulation environments are described and simulation results are shown in Section \ref{Simulation Environment and Results}. Finally, Section \ref{Conclusion} concludes the paper.

\section{LoRa} \label{LoRa}
\par LoRa is the name of the physical layer Radio/Chipset technology that provides wireless link solution for low power wide area networks. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). LoRa is registered trademark of Semtech Corporation. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently and incorrectly used for each other.

\subsection{LoRaWAN}
\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides inter-operability between different LoRa networks. LoRaWAN is based on pure ALOHA medium access. LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.

\par TODO: Add LoRaWAN network architecture figure here.

\par A typical LoRaWAN network consist of three network entities.
\subsubsection{End node}
Low power embedded device that communicates to gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios. Class A end nodes generates uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A by adding scheduled receive windows for downlink transmission. Receive window is synchronized using beacon packet transmitted by gateway. Class C end nodes extend Class A by keeping receive window open unless they are transmitting. This provides low latency donwlink communication but requires more power consumption. (CITE) In this paper, we focused on class A end devices since Class A provides lowest power consumption.

\subsubsection{Gateway}
Devices that receive/transmit packets coming from/to end nodes. A typical gateway can receive from multiple channel at the same time. Gateways are typically connected to power grid so power consumption of a gateway is insignificant most of the deployments.

\subsubsection{Network server}
Server that provides MAC layer processing. NS routes messages from application to end nodes, and vice versa.

\subsection{LoRa Modulation}
\par TODO
LoRa CSS modulation can spread the symbols.\\
TODO frequency and band

We invite readers to refer to \cite{AN1200.22} for further details.

\par Spreading factor is the ratio between symbol rate and chip rate. (CITE) The ratio between symbol and chip rate is equal to $2$\textsuperscript{SF}. SF can take values between 7 to 12.

and determines the LoRa symbol rate as: Rs = (BW/2SF). As the SF increases the symbol rate is lowered, thereby trading reduced data rate for increased range. For a bandwidth of 125 kHz, the PHY data rates range from 6835 to 365 b/s for SF7–SF12. The different SFs are orthogonal. This means that a LoRa gateway can receive multiple transmissions on different SFs simultaneously. Note that SF bits are mapped per LoRa symbol.

increasing the spread factor makes the signal more robust to noise, but decreases the data rate.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora0}
\caption{Different SF transmission ranges.}
\label{fig:problem1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora1}
\caption{Collision between nodes close to the GW.}
\label{fig:problem2}
\end{figure}

\section{Other Related Works} \label{Other Related Works}
The literature related to the work presented in this paper has started growing recently. LPWAN technologies and especially LoRa/LoRaWAN attract researchers attention lately. We listed some of these works which studying  LoRaWAN and LoRa spreading factor.

@@

\section{Proposed Technique} \label{Proposed Technique}
\par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF. Figure~\ref{fig:proposal_single_gw}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora2}
\caption{Collision between nodes close to the GW can be prevented by using higher SF.}
\label{fig:proposal_single_gw}
\end{figure}

@@

\section{Simulation Environment and Results} \label{Simulation Environment and Results}
A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs

@@

\section{Conclusion} \label{Conclusion}
TODO: Summary, future works
\cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
